---
title: "Data Analaysis with the Arctic shorebird csv files"
format: html
---


```{r, message=FALSE}
library(tidyverse)
library(lubridate)
library(RColorBrewer)
```

## Define path and filenames

```{r message=FALSE, warning = FALSE}
path_raw <- "data/raw_analysis"

sites_csv <- "site.csv"

snowcover_csv <- "snow_cover.csv"

nests_csv <- "ASDN_Bird_nests.csv"

eggs_csv <- "ASDN_Bird_eggs.csv"

species_csv <- "species.csv"

lemmings_csv <- "ASDN_Daily_pred_lemm.csv"

snowcover_data <- read_csv(file.path(path_raw, snowcover_csv))

# transform the date in date format
snowcover_data_fixed <- snowcover_data %>% 
  mutate(Date2 = as_date(Date)) 

# snowcover_data_fixed %>% 
#   filter(is.na(Date2)) %>% 
#   View()

snowcover_data_fixed <- snowcover_data_fixed %>% 
  mutate(Date = case_when(
    Date == "8&9 june 06" ~ "09-June-06", # update the date columns
    .default = Date
  )) %>% 
  mutate(Date2 = as_date(Date))

snowcover_data_fixed <- snowcover_data_fixed %>% 
  mutate(Date = case_when(
    Date == "8&9 june 06" ~ "09-June-06",
    .default = Date
  )) %>% 
  mutate(Date2 = as_date(Date, 
                         format = "%d-%b-%y")) # tell the funciton how to read the date from Date columns

#write.csv(snowcover_data_fixed, '/Users/lunacatalan/Documents/dev/eds213/bren-meds213-data-cleaning/data/processed/snow_cover.csv')

```

# Data Cleaning 

1. Clean the `Water_cover` column to transform it into the correct data type and respect expectations for a percentage

2. Clean the `Land_cover` column to transform it into the correct data type and respect expectations for a percentage

```{r message=FALSE, warning = FALSE}

snow_cover <- read_csv('/Users/lunacatalan/Documents/dev/eds213/bren-meds213-data-cleaning/data/processed/snow_cover.csv') %>% 
  mutate(Date = Date2) %>% 
  select(-Date2)

# snow_cover %>% 
#   count(Land_cover) %>% 
#   filter(is.na(as.numeric(Land_cover)))


# replace unknown and n/a values with NA's
clean_cover <- snow_cover %>% 
  
  # replace these values with NA
  mutate(Water_cover = ifelse(Water_cover == "unk", 
                              NA,
                              ifelse(Water_cover == "n/a",
                                     NA,
                                     Water_cover)),
         Land_cover = ifelse(Water_cover == "unk", 
                              NA,
                              ifelse(Land_cover == "n/a",
                                     NA,
                                     Land_cover))) %>% 
  mutate(Water_cover = as.numeric(Water_cover),
         Land_cover = as.numeric(Land_cover)) %>% 
  
  # remove the negative numbers - don't want to assume 100% water cover
  filter(Land_cover >= 0) %>% 
  mutate(id = `...1`) %>% 
  select(!`...1`)
  
  
```

Use the relationship between the three cover columns (Snow, Water, Land) to infer missing values where possible and recompute the Total_cover column

Create a dataframe that filters for the NA values in a total_check column
- if the snow, water, and land cover do not add up to total cover then set it to NA
- if they add together keep the values (which should be 100)

```{r warning=FALSE, message=FALSE}
total_check <- clean_cover %>% 
  
  #  if total_cover is equal to water + land then set to 0 since the math adds up
  mutate(Snow_cover = ifelse(Total_cover == (Water_cover + Land_cover),
                                    0,
                                    Snow_cover),
         
         # if all the columns equal 0, then set total cover to NA since there were no observations
         Total_cover = ifelse(Snow_cover == 0 & Water_cover == 0 & Land_cover == 0,
                              NA, # make NA because there were no valid observations
                              Total_cover)) %>% 
  
  # make a check column so that if the columns dont add up to total cover set the cover check to NA
  mutate(total_cover_check = ifelse(Total_cover == (Snow_cover + Water_cover + Land_cover),
                              Total_cover, 
                              NA)) %>% 
  filter(is.na(total_cover_check))
```

Check the NAs in 
- There are 622 rows where the NA values in total_check dont have observers...because of this I feel okay dropping the rows
- The total number of rows where snow, water, and land are empty is 1514...because of this I feel okay dropping these values
- This only leaves 10 rows where values seem to be mis-entered
- Because of these checks, I feel okay dropping all of the rows in the total_check dataframe from the `clean_cover`
```{r warning=FALSE, message=FALSE}
# check the number of observations where there are 0 for everything, and no observer
observer_check <- total_check %>% 
  filter(is.na(Observer))

# check the number of rows where all values equal 0
empty_check <- total_check %>% 
  filter(Snow_cover == 0 & Water_cover == 0 & Land_cover == 0)

clean_cover <- clean_cover %>% 
  filter(!id %in% total_check$id) # remove the rows that match the ones above

# write.csv(clean_cover, '/Users/lunacatalan/Documents/dev/eds213/bren-meds213-data-cleaning/data/processed/clean_cover.csv')
```


